### 2020-03-26 Markdown Documents/Markdown_Questions.Rmd

---
title: "Markdown/Github Questions for Matt Williamson"
author: "Matt Clark and Alex Killion"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Questions
1. Standard process for organizing citations using markdown.
2. Highlighting and commenting on text (analog to Word comment bubble)
3. How to work at the same time without creating merg conflicts
4. Should we be using .md files for writing, and only .rmd for anaylsis? Is there an advantage to .md?

### Answers
1.
2. I think you can do the comment stuff from the webpage using the 'inline comment' comment bit on the webpage
3. As far as writing at the same time, I'd use local branches (i.e., forks) to avoid that. So you create a fork called "IntroMC" and write that while Alex works on a fork called "mehtodsAK" when you're ready to pass it back, issue a pull request and merge te changes back to the master branch