### 2020-05-20 Markdown Documents/Manuscript.Rmd
Updated Plots

---
title: Estimating the Financial Impact of Wildfire Smoke on US National parks
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken?
- affiliation: Boise State University
  name: Laura? (undergrad who will continue working over summer)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: /Users/Matthewclark989/Documents/LatexPractice/templatedocs/svm_ms_template.tex
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
keywords: pandoc, r markdown, knitr
bibliography: /Users/Matthewclark989/Documents/SmokeProject/References.bib
thanks: If we want to thank anyone? Maybe Moji?.
abstract: National Parks are important. National Parks are low on funding which affects their ability to do their job. A key way they get this funding is through visitation. Wildfire smoke affects visitation greatly and is out of ther control. Wildfires are expected to increase in the Pacific West and Intermountain regions. Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in these regions. When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total. Wildfires and the smoke it produces is supposed to increase over the next 50 years. Assuming a linear trend in smoke, visitation overall, and park fees, we expect to see significant financial losses to the National Park Service. Our recomendation is that....NPS work with the forest service to prevent fires in areas close to parks?? Do they already do that??    
---
# 1. Introduction

This is how you add a reference (e.g., [@fisichelli_protected_2015; @fann_ch_2016])

Paragraphs:
1. Public lands are important
2. NPs are specifically high value
3. NPs are under funded. This under funding prevents them from doing those important jobs
4. NPs get a lot of their funding from visitation. Vis is influenced by a lot of things. One of these things that's particularly out of their control is smoke from adjacent forested areas.
5. Smoke is bad for recreation and should decrease vis.
6. Smoke events have been increasing over the last 50 years. Particularly, these increases have been in the west.
7. These smoke events are expected to worsen via climate change over the next 50 years.
8. Logically, smoke should influence visitation to parks. This decrease should affect income. We seek to assess this effect in a systematic way and backcast what the monetary effects might have been.
9. Given the trends we observe, we project out into the future what the effects might be over the next 50 years.

# 2. Methods
```{r include=FALSE, cache=TRUE}
library(tidyverse)
library(lme4)
library(rstanarm)
library(rstan)
library(zoo)
library(formattable)
load("~/SmokeProject/ModelObjects/feefit.rda")
load("~/SmokeProject/ModelObjects/residuals.rda")
load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
```
## 2.1. Data Collection

### 2.1.1. Visitor Data

### 2.1.2. Visitation Fees

We started with an incomplete dataset. Laura will write this part with information on collecting this data. 
```{r include=FALSE, cache=TRUE}
FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
```

We removed the NA's from this dataset built a linear model which estimates the fees based on the date and the park. 
```{r, echo=FALSE, cache=TRUE}
d<-na.omit(FEdat) #get rid of rows with any NA
```

$$
\begin{aligned}
y_{tp} &\sim \begin{cases}\operatorname{Pois}(\mu,\lambda)\\ \mu = B_{0}+ B_{1}*X_{1} +B_{2}*X_{2}
\end{cases}
\end{aligned}
$$

We operationalized this model in rstanarm

We then used this to predict what the fees would have been for every year we didn't have
```{r cache=TRUE, include=FALSE}
fillData<-FEdat[,1:2]

FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column

```

Our predicted fees correlate well with our observed fees that we have
```{r echo = FALSE, cache=TRUE}
xx<-na.omit(FEdat)
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                  plot.title = element_text( size=18, color="black",face="bold"),
                  axis.title.x = element_text( size=18),
                  axis.title.y = element_text( size=18),
                  axis.text=(element_text(color="black", size=14)),
                  legend.title = element_text(colour="black", size=18),
                  legend.text = element_text( size = 14))
xx%>%
  ggplot(.,aes(x=SingleVisitorEntryFee,y=PredictedFee))+
  geom_jitter(size=2, alpha=0.3)+
  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
```

### 2.1.3. Smoke Data

Alex to complete this section

## 2.2. Analysis

### 2.2.1. Smoke Effects Past
We subset the data to only include high season visitation for parks in the Inter-mountain and Pacific West regions. The regions projected to be most heavily affected by climate change.

```{r include=FALSE, cache=TRUE}
library(tidyverse)
library(lme4)
library(rstan)
Data<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
Data<-na.omit(Data)

#filter by high season
Data<-Data%>%filter( SeasType =="High")

#filter by smoke affected regions
Data<-Data%>%
  filter(Region %in% c("Intermountain","Pacific West"))

```

To account for the linear trends of visitation, we first calculated the trends, then found the residuals of each datapoint from the trend line.
$$
\begin{aligned}
Resid_{y} &= \begin{cases}y_{obs}-y_{trend}\\ y_{trend} \sim \operatorname{Norm}(\mu_{park},\sigma)\\ \mu_{park} = B_{0}+ B_{1}*X_{1}
\end{cases}
\end{aligned}
$$

```{r warning=FALSE, include=FALSE, cache=TRUE}
fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
Data$trendvis<-predict(fits,date=Data$date)
Data$VisDiff<-Data$RecreationVisits-Data$trendvis
dat<-Data
```

We then used a hierarchical modeling structure to understand these residuals as a result of the max observed smoke values in each park in each month.
$$
\begin{aligned}
Resid_{y} &\sim \begin{cases}\operatorname{Norm}(\mu,\sigma)\\ \mu = B_{0Park}+ B_{1Park}*X_{1Park} \\ \sigma\sim Norm(0,1) 
\end{cases}
\end{aligned}
$$

This was also operationalized in `rstanarm`
```{r eval=FALSE, cache=TRUE,include=FALSE}
fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)

```

### 2.2.2. Model Validation

```{r echo=FALSE, message=FALSE,cache=TRUE}
knitr::include_graphics(path.expand("~/SmokeProject/Train_Test.png"))
```

# 3. Results

## 3.1. Parameter Estimates
For all significant effects (90% C.I. doesn't overlap 0), we see a negative effect of smoke on the residual visitation to National Parks.

```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE}
PNames<-rev(unique(dat$ParkName))

pars<-plot(fit, regex_pars = "stdsmoke UnitCode",
     prob = 0.5, prob_outer = 0.9)

mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                plot.title = element_text( size=18, color="black",face="bold"),
                axis.title.x = element_text( size=18),
                axis.title.y = element_text( size=18),
                axis.text=(element_text(color="black", size=14)),
                legend.title = element_text(colour="black", size=18),
                legend.text = element_text( size = 14))
```

```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE, fig.height = 9, fig.width = 12}
pars+ggplot2::ggtitle("Posterior medians with 50% and 90% intervals")+
  ggplot2::scale_y_discrete(name ="Park",labels=PNames)+
  mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)
```

## 3.2. Backcasting
We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the visitation may have been given optimal air quality.

Here's an example from a few parks

```{r include=FALSE, cache=TRUE}

dat<-read.csv("~/SmokeProject/Data/FullDataWithMinSmokePredictions.csv")
dat$date<-zoo::as.yearmon(dat$date)
```

```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
library(scales)
p1<-predDat%>%filter(UnitCode == "GLAC")%>%
    ggplot(.,aes(x=date))+
    geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
    geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
    geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
    scale_color_identity(name = "",
                         breaks = c("black", "#1f78b4"),
                         labels = c("Observed", "Median Estimate Minimum Smoke"),
                         guide = "legend")+ggtitle("Glacier National Park")+theme(legend.position="top")+
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.line.x = element_blank())

p2<-predDat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
    summarise(meansmoke=mean(Smoke))%>%
    ggplot(.,aes(x=Year,y=meansmoke))+
    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
    geom_point(aes(color =(meansmoke)),size=5)+
    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
                          labels=c("Above","Below")
    ) + 
    #guides(color = FALSE) +
    guides(alpha = FALSE) + 
    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 9e-10))+mytheme

gA <- ggplotGrob(p1)
gB <- ggplotGrob(p2)
grid::grid.newpage()
grid::grid.draw(rbind(gA, gB))
```

```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
library(scales)
p1<-predDat%>%filter(UnitCode == "YELL")%>%
    ggplot(.,aes(x=date))+
    geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
    geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
    geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
    scale_color_identity(name = "",
                         breaks = c("black", "#1f78b4"),
                         labels = c("Observed", "Median Estimate Minimum Smoke"),
                         guide = "legend")+ggtitle("Yellowstone National Park")+theme(legend.position="top")+
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.line.x = element_blank())

p2<-predDat%>%filter(UnitCode == "YELL")%>%group_by(Year)%>%
    summarise(meansmoke=mean(Smoke))%>%
    ggplot(.,aes(x=Year,y=meansmoke))+
    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
    geom_point(aes(color =(meansmoke)),size=5)+
    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
                          labels=c("Above","Below")
    ) + 
    #guides(color = FALSE) +
    guides(alpha = FALSE) + 
    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 9e-10))+mytheme

gA <- ggplotGrob(p1)
gB <- ggplotGrob(p2)
grid::grid.newpage()
grid::grid.draw(rbind(gA, gB))
```

We can apply these visitation estimates to the fee data/predictions also. This gives us the following table.
```{r echo=FALSE, cache=TRUE}
fitint<-as.data.frame(posterior_interval(fit))
names(fitint)<-c("CI5","CI95")
fitint$Par<-row.names(fitint)
fitsig<-fitint[grep("stdsmoke UnitCode",row.names(fitint)),]%>% 
  filter(CI5<0 & CI95<0)

fitsig$Par<-substr(fitsig$Par,21,24)

#predicted vis fees 
FEdat<-read.csv("~/SmokeProject/Data/ParkFees.csv")

#get the visitation observed and predicted yearly totals (same scale as fee data)
VEstDat<-dat%>%group_by(UnitCode,Year)%>%
  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke50CI))%>%
  filter(UnitCode %in% fitsig$Par)

#only include significant parks and years we predicted for
FEdat<-FEdat%>%filter(UnitCode%in%fitsig$Par)

#merge the fee and visitation data
financialData<-merge(FEdat, VEstDat, by = c("UnitCode","Year"))

#multiply the fees and visitation (predicted and observed)
financialData<-financialData%>%mutate(incomeObserved= round(SingleVisitorFee*RecVisits,digits = 2),incomePred= round(SingleVisitorFee*PredVisits,digits=2))

#get the dollar estimates for smoke loss for each park)
summaryFinancialLoss<-financialData %>% group_by(UnitCode)%>%summarise(totrec=sum(incomeObserved),totpred=sum(incomePred))%>%
  mutate(IncomeDif=totrec-totpred,IncomePrecDiff=round((1-(totrec/totpred))*100,digits=2))

summaryFinancialLoss$ParkName<-unique(dat[dat$UnitCode %in% summaryFinancialLoss$UnitCode,]$ParkName)

summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
summaryFinancialLoss$IncomeDif<-currency(summaryFinancialLoss$IncomeDif)
names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
formattable::formattable(summaryFinancialLoss)
```
## 3.3. Model Validation
To assess the validity of our model, we used a k-fold cross validation technique.

```{r echo=FALSE, message=FALSE,cache=TRUE}
library(tidyverse)
library(rstanarm)
#pull out last 5 years
train1<-dat%>%filter(!Year %in% c("2018","2017","2016","2015","2014"))
test1<-dat%>%filter(Year %in% c("2018","2017","2016","2015","2014"))

#fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train1,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain,file="CVResidualMod.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod.rds")

preds1<-apply(posterior_predict(fitTrain,test1),2,median)
reals1<-test1$VisDiff

## 2013:2009

train2<-dat%>%filter(!Year %in% c("2013","2012","2011","2010","2009"))
test2<-dat%>%filter(Year %in% c("2013","2012","2011","2010","2009"))
#fitTrain2<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train2,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain2,file="CVResidualMod2.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")

preds2<-apply(posterior_predict(fitTrain2,test2),2,median)
reals2<-test2$VisDiff

# 2008:2004
train3<-dat%>%filter(!Year %in% c("2008","2007","2006","2005","2004"))
test3<-dat%>%filter(Year %in% c("2008","2007","2006","2005","2004"))
#fitTrain3<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train3,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain3,file="CVResidualMod3.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod3.rds")

preds3<-apply(posterior_predict(fitTrain3,test3),2,median)
reals3<-test3$VisDiff

# 2003:1999
train4<-dat%>%filter(!Year %in% c("2003","2002","2001","2000","1999"))
test4<-dat%>%filter(Year %in% c("2003","2002","2001","2000","1999"))
#fitTrain4<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train4,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain4,file="CVResidualMod4.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod4.rds")

preds4<-apply(posterior_predict(fitTrain4,test4),2,median)
reals4<-test4$VisDiff

# 1998:1994
train5<-dat%>%filter(!Year %in% c("1998","1997","1996","1995","1994"))
test5<-dat%>%filter(Year %in% c("1998","1997","1996","1995","1994"))
#fitTrain5<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train5,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain5,file="CVResidualMod5.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod5.rds")

preds5<-apply(posterior_predict(fitTrain5,test5),2,median)
reals5<-test5$VisDiff

# 1993:1989
train6<-dat%>%filter(!Year %in% c("1993","1992","1991","1990","1989"))
test6<-dat%>%filter(Year %in% c("1993","1992","1991","1990","1989"))
#fitTrain6<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train6,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain6,file="CVResidualMod6.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod6.rds")

preds6<-apply(posterior_predict(fitTrain6,test6),2,median)
reals6<-test6$VisDiff

# 1988:1980
train7<-dat%>%filter(!Year %in% c("1988","1987","1986","1985","1984","1983","1982","1981","1980"))
test7<-dat%>%filter(Year %in% c("1988","1987","1986","1985","1984","1983","1982","1981","1980"))
#fitTrain7<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train7,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain7,file="CVResidualMod7.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod7.rds")

preds7<-apply(posterior_predict(fitTrain7,test7),2,median)
reals7<-test7$VisDiff

#combine them all
preds<-c(preds1,preds2,preds3,preds4,preds5,preds6,preds7)
reals<-c(reals1,reals2,reals3,reals4,reals5,reals6,reals7)

d<-as.data.frame(cbind(as.vector(preds),reals))
names(d)<-c("preds","reals")
mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                plot.title = element_text( size=18, color="black",face="bold"),
                axis.title.x = element_text( size=18),
                axis.title.y = element_text( size=18),
                axis.text=(element_text(color="black", size=14)),
                legend.title = element_text(colour="black", size=18),
                legend.text = element_text( size = 14))
d%>%
  ggplot(.,aes(x=reals,y=preds))+
  geom_point(size=2.5, alpha=0.2)+
  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
  geom_text(aes(x=3e+05,y=-2e+05),size=10,label=paste("cor =",round(cor(d$preds,d$reals),digits=2)))+
  ggtitle("CV Results")+theme_classic()+mytheme+scale_x_continuous("Observed Residual Visitation", labels = scales::comma)+
  scale_y_continuous(name="Predicted Residual Visitation", labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))

```

```{r echo=FALSE, message=FALSE,cache=TRUE}
x<-data.frame(Model=c(1:7), 
              "TimePeriod"=c("1980 - 1988",
                                         "1989 - 1993",
                                         "1994 - 1998",
                                         "1999 - 2003",
                           "2004 - 2008",
                           "2009 - 2013",
                           "2014 - 2018"), 
              Correlation=c(cor(as.vector(preds1),reals1),
                                                   cor(as.vector(preds2),reals2),
                                                   cor(as.vector(preds3),reals3),
                                                   cor(as.vector(preds4),reals4),
                                                   cor(as.vector(preds5),reals5),
                                                   cor(as.vector(preds6),reals6),
                                                   cor(as.vector(preds7),reals7)))

x$r2<-round(x$Correlation^2,digits=2)
x$Correlation<-round(x$Correlation,digits=2)
names(x)<-c("Model","Time Period", "Correlation", "R^2")
formattable::formattable(x)
```

# 4. Discussion

Discussion here