### 2020-05-20 Markdown Documents/Manuscript.Rmd
Rough draft of intro done

---
title: Estimating the Financial Impact of Wildfire Smoke on US National parks
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken?
- affiliation: Boise State University
  name: Laura? (undergrad who will continue working over summer)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: /Users/Matthewclark989/Documents/LatexPractice/templatedocs/svm_ms_template.tex
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
keywords: pandoc, r markdown, knitr
bibliography: /Users/Matthewclark989/Documents/SmokeProject/References.bib
thanks: If we want to thank anyone? Maybe Moji?.
abstract: National Parks are important. National Parks are low on funding which affects their ability to do their job. A key way they get this funding is through visitation. Wildfire smoke affects visitation greatly and is out of ther control. Wildfires are expected to increase in the Pacific West and Intermountain regions. Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in these regions. When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total. Wildfires and the smoke it produces is supposed to increase over the next 50 years. Assuming a linear trend in smoke, visitation overall, and park fees, we expect to see significant financial losses to the National Park Service. Our recomendation is that....NPS work with the forest service to prevent fires in areas close to parks?? Do they already do that??    
---
# 1. Introduction

Twenty eight percent of the land area of the United States is designated as public land [@vincent_federal_2020]. These public lands are the critical backdrop for the majority wildlife conservation and outdoor recreation in the country [@romagosa_inside_2015; @leemans_millennium_2003; @dudley_national_2011]. In a 2016 study, 93% of respondents indicated that they believe public open spaces and recreational areas should be protected, regardless of whether or not they visited them personally [@haefele_total_2016]. The crown jewel of the United States public lands system is the National Park Service. The national park system alone received approximately 318 million recreational visitors and generated over $40 billion in economic output to local economies in 2018 [@thomas_2018_2018].

National parks undeniably represent a critical infrastructure for supporting ecological integrity and outdoor recreation in the United States, but have recently drawn public attention for considerable under funding [@simmonds_crisis_2018;@duncan_opinion_2016]. Most notably for the $12 billion repair backlog for basic maintenance and resource protection [@nps_nps_2018]. In addition to necessary repairs, the National Park Service reports a 14% reduction in staff funding over the last 5 years [@doi_fiscal_2019]. While the majority of National Park Service funding comes from congressional appropriations, recent budgetary restrictions have put pressure on parks to gather an increasing amount of their budget from recreational visitor fees such as admissions and camping. The income from these fees has steadily increased over the last 10 years as park visitation and visitor fees have both surged [@gao_national_2016].

As national parks become increasingly dependent on visitor fees in order to provide the suite of social and ecological services communities depend on, it is of paralleled importance to understand what drives fluctuation in recreational visits. Research has shown that economic considerations (e.g. park fees) are considerably less influential on park visitation than are recreational opportunities [@neuvonen_visits_2010;@stevens_declining_2014]. Most factors which influence the recreational accessibility of a given park are either fixed (natural resources within the park area) or within the control of park management (permitting, trail building, etc.). Wildfire smoke however, shows considerable yearly variation and is largely produced by adjacent public lands and therefore out of the jurisdiction of park management [@dombeck_wildfire_2004;@collins_regional_2006].

Wildfire smoke is a mixture of many gasses aerated by biomass combustion across a landscape [@naeher_woodsmoke_2007]. While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 μm in diameter (PM2.5) is particularly harmful to human health, causing between 260,000 - 600,000 global deaths annually [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007]. Health concerns regarding wildfire smoke, specifically PM2.5, regurlarly drive calls for individuals to reduce outdoor recreation behavior [@oneill_summary_2013].

Wildfires and the smoke they produce have been increasing in the United States over the last 50 years, specifically in the American West and Inter-Mountain regions [@kitzberger_contingent_2007;@westerling_increasing_2016]. These regions are also projected to experience increased wildfire frequency and severity over the next 50 years as a result of climate change [@harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015]. Anecdotally, it is known that wildfire smoke has impacted the recreation behavior of individuals in these areas and decreased visitation revenue to the national parks in both regions. This assumption however has not yet been demonstrated empirically.

The primary goal of this project is to assess how visitation to national parks in the Inter-Mountain and Pacific regions has systematically varied with surface level concentrations of PM2.5 resulting from wildfire smoke from 1980 - 2018. Given the variability in recreational fees between parks and over time, the secondary goal of this project is to produce order of magnitude estimates of the financial losses that wildfire smoke has caused  park units during the same time period. The intention of this research is not to perform an economic analysis of smoke effects on each park specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and may be in the future.

# 2. Methods
```{r include=FALSE, cache=TRUE}
library(tidyverse)
library(lme4)
library(rstanarm)
library(rstan)
library(zoo)
library(formattable)
load("~/SmokeProject/ModelObjects/feefit.rda")
load("~/SmokeProject/ModelObjects/residuals.rda")
load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
```
## 2.1. Data Collection

### 2.1.1. Visitor Data

### 2.1.2. Visitation Fees

We started with an incomplete dataset. Laura will write this part with information on collecting this data. 
```{r include=FALSE, cache=TRUE}
FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
```

We removed the NA's from this dataset built a linear model which estimates the fees based on the date and the park. 
```{r, echo=FALSE, cache=TRUE}
d<-na.omit(FEdat) #get rid of rows with any NA
```

$$
\begin{aligned}
y_{tp} &\sim \begin{cases}\operatorname{Pois}(\mu,\lambda)\\ \mu = B_{0}+ B_{1}*X_{1} +B_{2}*X_{2}
\end{cases}
\end{aligned}
$$

We operationalized this model in rstanarm

We then used this to predict what the fees would have been for every year we didn't have
```{r cache=TRUE, include=FALSE}
fillData<-FEdat[,1:2]

FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column

```

Our predicted fees correlate well with our observed fees that we have
```{r echo = FALSE, cache=TRUE}
xx<-na.omit(FEdat)
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                  plot.title = element_text( size=18, color="black",face="bold"),
                  axis.title.x = element_text( size=18),
                  axis.title.y = element_text( size=18),
                  axis.text=(element_text(color="black", size=14)),
                  legend.title = element_text(colour="black", size=18),
                  legend.text = element_text( size = 14))
xx%>%
  ggplot(.,aes(x=SingleVisitorEntryFee,y=PredictedFee))+
  geom_jitter(size=2, alpha=0.3)+
  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
```

### 2.1.3. Smoke Data

Alex to complete this section

## 2.2. Analysis

### 2.2.1. Smoke Effects Past
We subset the data to only include high season visitation for parks in the Inter-mountain and Pacific West regions. The regions projected to be most heavily affected by climate change.

```{r include=FALSE, cache=TRUE}
library(tidyverse)
library(lme4)
library(rstan)
Data<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
Data<-na.omit(Data)

#filter by high season
Data<-Data%>%filter( SeasType =="High")

#filter by smoke affected regions
Data<-Data%>%
  filter(Region %in% c("Intermountain","Pacific West"))

```

To account for the linear trends of visitation, we first calculated the trends, then found the residuals of each datapoint from the trend line.
$$
\begin{aligned}
Resid_{y} &= \begin{cases}y_{obs}-y_{trend}\\ y_{trend} \sim \operatorname{Norm}(\mu_{park},\sigma)\\ \mu_{park} = B_{0}+ B_{1}*X_{1}
\end{cases}
\end{aligned}
$$

```{r warning=FALSE, include=FALSE, cache=TRUE}
fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
Data$trendvis<-predict(fits,date=Data$date)
Data$VisDiff<-Data$RecreationVisits-Data$trendvis
dat<-Data
```

We then used a hierarchical modeling structure to understand these residuals as a result of the max observed smoke values in each park in each month.
$$
\begin{aligned}
Resid_{y} &\sim \begin{cases}\operatorname{Norm}(\mu,\sigma)\\ \mu = B_{0Park}+ B_{1Park}*X_{1Park} \\ \sigma\sim Norm(0,1) 
\end{cases}
\end{aligned}
$$

This was also operationalized in `rstanarm`
```{r eval=FALSE, cache=TRUE,include=FALSE}
fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)

```

### 2.2.2. Model Validation

```{r echo=FALSE, message=FALSE,cache=TRUE}
knitr::include_graphics(path.expand("~/SmokeProject/Train_Test.png"))
```

# 3. Results

## 3.1. Parameter Estimates
For all significant effects (90% C.I. doesn't overlap 0), we see a negative effect of smoke on the residual visitation to National Parks.

```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE}
PNames<-unique(dat$ParkName)

pp<-as.data.frame(posterior_interval(fit))
limits<-row.names(pp[grep("stdsmoke UnitCode",row.names(pp)),])

pars<-plot(fit, regex_pars = "stdsmoke UnitCode",
           prob = 0.5, prob_outer = 0.9)

mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                plot.title = element_text( size=18, color="black",face="bold"),
                axis.title.x = element_text( size=18),
                axis.title.y = element_text( size=18),
                axis.text=(element_text(color="black", size=14)),
                legend.title = element_text(colour="black", size=18),
                legend.text = element_text( size = 14))

```

```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE, fig.height = 9, fig.width = 12}
pars+ggplot2::ggtitle("Posterior medians with 50% and 90% intervals")+
  ggplot2::scale_y_discrete(name ="Park",labels=rev(PNames),limits=rev(limits))+
  mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)
```

## 3.2. Backcasting
We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the visitation may have been given optimal air quality.

Here's an example from a few parks

```{r include=FALSE, cache=TRUE}

dat<-read.csv("~/SmokeProject/Data/FullDataWithMinSmokePredictions.csv")
dat$date<-zoo::as.yearmon(dat$date)
```

```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
library(scales)
p1<-dat%>%filter(UnitCode == "GLAC")%>%
    ggplot(.,aes(x=date))+
    geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
    geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
    geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
    scale_color_identity(name = "",
                         breaks = c("black", "#1f78b4"),
                         labels = c("Observed", "Median Estimate Minimum Smoke"),
                         guide = "legend")+ggtitle("Glacier National Park")+theme(legend.position="top")+
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.line.x = element_blank())

p2<-dat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
    summarise(meansmoke=mean(Smoke))%>%
    ggplot(.,aes(x=Year,y=meansmoke))+
    geom_segment(aes(xend = Year, yend = 2.5e-10,color =(meansmoke)),size=1.5) +
    geom_point(aes(color =(meansmoke)),size=5)+
    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
                          labels=c("Above","Below")
    ) + 
    #guides(color = FALSE) +
    guides(alpha = FALSE) + 
    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(2.5e-10, 1.25e-9))+mytheme

gA <- ggplotGrob(p1)
gB <- ggplotGrob(p2)
grid::grid.newpage()
grid::grid.draw(rbind(gA, gB))
```

```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
library(scales)
p1<-dat%>%filter(UnitCode == "OLYM")%>%
    ggplot(.,aes(x=date))+
    geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
    geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
    geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
    scale_color_identity(name = "",
                         breaks = c("black", "#1f78b4"),
                         labels = c("Observed", "Median Estimate Minimum Smoke"),
                         guide = "legend")+ggtitle("Olympic National Park")+theme(legend.position="top")+
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.line.x = element_blank())

p2<-dat%>%filter(UnitCode == "OLYM")%>%group_by(Year)%>%
    summarise(meansmoke=mean(Smoke))%>%
    ggplot(.,aes(x=Year,y=meansmoke))+
    geom_segment(aes(xend = Year, yend = 1e-10,color =(meansmoke)),size=1.5) +
    geom_point(aes(color =(meansmoke)),size=5)+
    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
                          labels=c("Above","Below")
    ) + 
    #guides(color = FALSE) +
    guides(alpha = FALSE) + 
    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(1e-10, 9e-10))+mytheme

gA <- ggplotGrob(p1)
gB <- ggplotGrob(p2)
grid::grid.newpage()
grid::grid.draw(rbind(gA, gB))
```

We can apply these visitation estimates to the fee data/predictions also. This gives us the following table.
```{r echo=FALSE, cache=TRUE}
fitint<-as.data.frame(posterior_interval(fit))
names(fitint)<-c("CI5","CI95")
fitint$Par<-row.names(fitint)
fitsig<-fitint[grep("stdsmoke UnitCode",row.names(fitint)),]%>% 
  filter(CI5<0 & CI95<0)

fitsig$Par<-substr(fitsig$Par,21,24)

#predicted vis fees 
FEdat<-read.csv("~/SmokeProject/Data/ParkFees.csv")

#get the visitation observed and predicted yearly totals (same scale as fee data)
VEstDat<-dat%>%group_by(UnitCode,Year)%>%
  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke50CI))%>%
  filter(UnitCode %in% fitsig$Par)

#only include significant parks and years we predicted for
FEdat<-FEdat%>%filter(UnitCode%in%fitsig$Par)

#merge the fee and visitation data
financialData<-merge(FEdat, VEstDat, by = c("UnitCode","Year"))

#multiply the fees and visitation (predicted and observed)
financialData<-financialData%>%mutate(incomeObserved= round(SingleVisitorFee*RecVisits,digits = 2),incomePred= round(SingleVisitorFee*PredVisits,digits=2))

#get the dollar estimates for smoke loss for each park)
summaryFinancialLoss<-financialData %>% group_by(UnitCode)%>%summarise(totrec=sum(incomeObserved),totpred=sum(incomePred))%>%
  mutate(IncomeDif=totrec-totpred,IncomePrecDiff=round((1-(totrec/totpred))*100,digits=2))

summaryFinancialLoss$ParkName<-unique(dat[dat$UnitCode %in% summaryFinancialLoss$UnitCode,]$ParkName)

summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
summaryFinancialLoss$IncomeDif<-currency(summaryFinancialLoss$IncomeDif)
names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
formattable::formattable(summaryFinancialLoss)
```
## 3.3. Model Validation
To assess the validity of our model, we used a k-fold cross validation technique.

```{r echo=FALSE, message=FALSE,cache=TRUE}
library(tidyverse)
library(rstanarm)
#pull out last 5 years
train1<-dat%>%filter(!Year %in% c("2018","2017","2016","2015","2014"))
test1<-dat%>%filter(Year %in% c("2018","2017","2016","2015","2014"))

#fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train1,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain,file="CVResidualMod.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod.rds")

preds1<-apply(posterior_predict(fitTrain,test1),2,median)
reals1<-test1$VisDiff

## 2013:2009

train2<-dat%>%filter(!Year %in% c("2013","2012","2011","2010","2009"))
test2<-dat%>%filter(Year %in% c("2013","2012","2011","2010","2009"))
#fitTrain2<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train2,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain2,file="CVResidualMod2.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")

preds2<-apply(posterior_predict(fitTrain2,test2),2,median)
reals2<-test2$VisDiff

# 2008:2004
train3<-dat%>%filter(!Year %in% c("2008","2007","2006","2005","2004"))
test3<-dat%>%filter(Year %in% c("2008","2007","2006","2005","2004"))
#fitTrain3<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train3,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain3,file="CVResidualMod3.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod3.rds")

preds3<-apply(posterior_predict(fitTrain3,test3),2,median)
reals3<-test3$VisDiff

# 2003:1999
train4<-dat%>%filter(!Year %in% c("2003","2002","2001","2000","1999"))
test4<-dat%>%filter(Year %in% c("2003","2002","2001","2000","1999"))
#fitTrain4<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train4,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain4,file="CVResidualMod4.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod4.rds")

preds4<-apply(posterior_predict(fitTrain4,test4),2,median)
reals4<-test4$VisDiff

# 1998:1994
train5<-dat%>%filter(!Year %in% c("1998","1997","1996","1995","1994"))
test5<-dat%>%filter(Year %in% c("1998","1997","1996","1995","1994"))
#fitTrain5<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train5,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain5,file="CVResidualMod5.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod5.rds")

preds5<-apply(posterior_predict(fitTrain5,test5),2,median)
reals5<-test5$VisDiff

# 1993:1989
train6<-dat%>%filter(!Year %in% c("1993","1992","1991","1990","1989"))
test6<-dat%>%filter(Year %in% c("1993","1992","1991","1990","1989"))
#fitTrain6<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train6,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain6,file="CVResidualMod6.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod6.rds")

preds6<-apply(posterior_predict(fitTrain6,test6),2,median)
reals6<-test6$VisDiff

# 1988:1980
train7<-dat%>%filter(!Year %in% c("1988","1987","1986","1985","1984","1983","1982","1981","1980"))
test7<-dat%>%filter(Year %in% c("1988","1987","1986","1985","1984","1983","1982","1981","1980"))
#fitTrain7<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train7,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
#save(fitTrain7,file="CVResidualMod7.rds")
load("~/SmokeProject/ModelObjects/CVResidualMod7.rds")

preds7<-apply(posterior_predict(fitTrain7,test7),2,median)
reals7<-test7$VisDiff

#combine them all
preds<-c(preds1,preds2,preds3,preds4,preds5,preds6,preds7)
reals<-c(reals1,reals2,reals3,reals4,reals5,reals6,reals7)

d<-as.data.frame(cbind(as.vector(preds),reals))
names(d)<-c("preds","reals")
mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                plot.title = element_text( size=18, color="black",face="bold"),
                axis.title.x = element_text( size=18),
                axis.title.y = element_text( size=18),
                axis.text=(element_text(color="black", size=14)),
                legend.title = element_text(colour="black", size=18),
                legend.text = element_text( size = 14))
d%>%
  ggplot(.,aes(x=reals,y=preds))+
  geom_point(size=2.5, alpha=0.2)+
  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
  geom_text(aes(x=3e+05,y=-2e+05),size=10,label=paste("cor =",round(cor(d$preds,d$reals),digits=2)))+
  ggtitle("CV Results")+theme_classic()+mytheme+scale_x_continuous("Observed Residual Visitation", labels = scales::comma)+
  scale_y_continuous(name="Predicted Residual Visitation", labels = scales::comma)+
  theme(plot.title = element_text(hjust = 0.5))

```

```{r echo=FALSE, message=FALSE,cache=TRUE}
x<-data.frame(Model=c(1:7), 
              "TimePeriod"=c("1980 - 1988",
                                         "1989 - 1993",
                                         "1994 - 1998",
                                         "1999 - 2003",
                           "2004 - 2008",
                           "2009 - 2013",
                           "2014 - 2018"), 
              Correlation=c(cor(as.vector(preds1),reals1),
                                                   cor(as.vector(preds2),reals2),
                                                   cor(as.vector(preds3),reals3),
                                                   cor(as.vector(preds4),reals4),
                                                   cor(as.vector(preds5),reals5),
                                                   cor(as.vector(preds6),reals6),
                                                   cor(as.vector(preds7),reals7)))

x$r2<-round(x$Correlation^2,digits=2)
x$Correlation<-round(x$Correlation,digits=2)
names(x)<-c("Model","Time Period", "Correlation", "R^2")
formattable::formattable(x)
```

# 4. Discussion

Discussion here