### 2020-07-01 Rcode/Carbon_Analysis.Rmd
update

---
title: 'MERRA-2: Carbon Analysis'
author: "Alex Killion"
date: "6/25/2020"
output: html_document
---

```{r}
https://www.r-bloggers.com/a-netcdf-4-in-r-cheatsheet/

url = "https://goldsmr5.gesdisc.eosdis.nasa.gov/opendap/MERRA2/M2I3NVAER.5.12.4/1980/01/MERRA2_100.inst3_3d_aer_Nv.19800101.nc4"
data = nc_open(url)

#have to delete duplicate ".nc4" in url
url2 = "https://goldsmr5.gesdisc.eosdis.nasa.gov/opendap/MERRA2/M2I3NVAER.5.12.4/2018/07/MERRA2_400.inst3_3d_aer_Nv.20180708.nc4?BCPHILIC[0:7][0:71][225:277][75:177],OCPHILIC[0:7][0:71][225:277][75:177],BCPHOBIC[0:7][0:71][225:277][75:177],OCPHOBIC[0:7][0:71][225:277][75:177],lat[225:277],time[0:7],lon[75:177],lev"
data2 = nc_open(url2)

#time, level, lat, long
get<-ncvar_get(data2, varid = c("BCPHOBIC","OCPHILIC","BCPHILIC","OCPHOBIC"))

#extract all 4 variables

#average by month

#make spatial

#clip with parks

```

```