### 2020-07-16 Rcode/Merra2.Rmd

---
title: "Merra2"
author: "Alex Killion"
date: "11/26/2019"
output: html_document
---

Packages
```{r}
library(ncdf4)
library(chron)
library(lattice)
library(RColorBrewer)

library(raster)
library(rgdal)
library(exactextractr)
library(readr)
library(plyr)
library(purrr)
```

*For MATRIX*
Open netCDF
```{r}
# set path and filename
ncpath <- "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/Merra2/"
ncname <- "MERRA2_100.tavgM_2d_aer_Nx.198001.SUB"
ncfname <- paste(ncpath, ncname, ".nc", sep="")
dname <- "DUSMASS25"

ncin <- nc_open(ncfname)

# get longitude and latitude and time
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
print(c(nlon,nlat))

time <- ncvar_get(ncin,"time")

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time) #dim() tells you number of timesteps
nt
```

Extract a Variable
```{r}
# get temperature
tmp_array <- ncvar_get(ncin,dname) #dname in 1st chunk needs to be correct variable name
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(tmp_array)

#close file
nc_close(ncfname)
```

Convert Time
```{r}
# convert time -- split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])
chron(time,origin=c(tmonth, tday, tyear))
```

Replace with NA
```{r}
# replace netCDF fill values with NA's
tmp_array[tmp_array==fillvalue$value] <- NA
```

Map
```{r}
# quick map
image(lon,lat,tmp_array, col=rev(brewer.pal(10,"RdBu")))
```

*For Raster*
Stack netCDF
```{r}
filenames <- list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/Merra2/",
                        pattern='*.nc',full.names=TRUE)
#if you want to stack
#smoke <- raster::stack(filenames, varname="DUSMASS25")

```

Import NPS Boundaries
```{r}
#Data From - https://public-nps.opendata.arcgis.com/datasets/national-park-service-park-unit-boundaries

nps<-readOGR("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/NPS_poly/NPS_poly.shp")

#select only polygons designated as National Parks
nps<-nps[nps@data$UNIT_TYPE=="National Park",]

#remove National Parks outside our Regions
nps<-nps[nps@data$REGION=="PW"|nps@data$REGION=="IM",]
nps<-nps[nps@data$UNIT_CODE!="NPSA",] #Samoa
nps<-nps[nps@data$STATE!="HI",]

#reformat to use park codes later
nps_df<-as.data.frame(nps)
nps_df<-as.matrix(nps_df$UNIT_CODE)

park_codes<-as.data.frame(nps_df)
names(park_codes)<-"park"

```

New data mean in Parks
```{r}

poly<-sf::st_as_sfc(nps) #converts to new type of poly

fileNames<-list.files("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT",full.names = TRUE)

years<-c(1980,2004)  #all the years
months<-c("01","10") #all the months
comb<-expand.grid(years,months) #just to get the number of unique combinations
list_all<-list() #create an empty list to fill with the file names

for (i in 1:nrow(comb)){ #creates the number of list objects
  year<-comb[i,]$Var1
  month<-comb[i,]$Var2
      x<-(paste0(year,month))
      list_all[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
      names(list_all)[i]<-paste0("x",year,month) #name the list objects
}

# list_all<- list_all[c(1,4)] You will need to use this with our sample data folder

empty_list<-list()

for (i in 1:length (list_all)) {
    for (j in 1:length (list_all[[i]])){
 bcphilic<-mean(brick(list_all[[i]][[j]], varname="BCPHILIC", level=72)) #one day - first bracket is month, 2nd is day, within a day is 8 layers by hour - the "level" dimension refers to veritcal layers where 72=surface and 1=top of atmosphere
bcphobic<-mean(brick(list_all[[i]][[j]], varname="BCPHOBIC", level=72))
ocphilic<-mean(brick(list_all[[i]][[j]], varname="OCPHILIC", level=72))
ocphobic<-mean(brick(list_all[[i]][[j]], varname="OCPHOBIC", level=72))

all<-bcphilic+bcphobic+ocphilic+ocphobic

empty_list[[j]]<-exact_extract(all, poly, 'mean') #intersects boundary

    }

out<-as.data.frame(empty_list)
out_mean<-as.data.frame(rowMeans(out)) #take mean of the assign list above

#write csv
  n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
  colnames(out_mean)[1] <- n #give values column name based on month
   name<-paste0("/Users/killion/output_means/",n) #create empty folder to save to
    newname<-paste0(name, ".csv")
write.csv(out_mean, newname, row.names = FALSE) #save as csv

  }

#loads those .csv and makes into single df
list_dates = list.files(path="/Users/killion/output_means/", pattern="*.csv", full.names=TRUE) #same folder you saved above output

final_means<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_means<-data.frame(park=park_codes, final_means)

#save
write.csv(final_means, "/Users/killion/final_means2.csv", row.names = FALSE)

```

Mean PM2.5 in Parks
```{r}

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

filenames <- list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/Merra2/",
                        pattern='*.nc',full.names=TRUE)

#loop to calculate polygons means and save vector
  for (i in 1:length (filenames)) {
    raster<-raster(filenames[i], varname="DUSMASS25")#convert smoke data to raster
    mean<-exact_extract(raster, poly, 'mean') #extracts mean for each park. Uses cells that intersect polygons.
   date<-str_sub(filenames[i], -13,-8) #strips down filename to just date
  mean<-as.data.frame(mean) #convert to df
  names(mean)<-date #rename column with date
  date_file<-paste0("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_mean/",date)
   newname<-paste0(date_file, ".csv")
write.csv(mean, newname, row.names = FALSE) #save as csv
}

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_mean/", pattern="*.csv", full.names=TRUE)

final_means<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_means<-data.frame(park=park_codes, final_means)

#save
write.csv(final_means, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_means.csv", row.names = FALSE)

```

Min PM2.5 in Parks
```{r}

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

#loop to calculate polygons means and save vector
  for (i in 1:length (filenames)) {
    raster<-raster(filenames[i], varname="DUSMASS25")#convert smoke data to raster
    min<-exact_extract(raster, poly, 'min') #extracts mean for each park. Uses cells that intersect polygons.
   date<-str_sub(filenames[i], -13,-8) #strips down filename to just date
  min<-as.data.frame(min) #convert to df
  names(min)<-date #rename column with date
  date_file<-paste0("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_min/",date)
   newname<-paste0(date_file, ".csv")
write.csv(min, newname, row.names = FALSE) #save as csv
}

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_min/", pattern="*.csv", full.names=TRUE)

final_mins<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_mins<-data.frame(park=park_codes, final_mins)

#save
write.csv(final_mins, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_mins.csv", row.names = FALSE)

```

Max PM2.5 in Parks
```{r}

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

#loop to calculate polygons means and save vector
  for (i in 1:length (filenames)) {
    raster<-raster(filenames[i], varname="DUSMASS25")#convert smoke data to raster
    max<-exact_extract(raster, poly, 'max') #extracts mean for each park. Uses cells that intersect polygons.
   date<-str_sub(filenames[i], -13,-8) #strips down filename to just date
  max<-as.data.frame(max) #convert to df
  names(max)<-date #rename column with date
  date_file<-paste0("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_max/",date)
   newname<-paste0(date_file, ".csv")
write.csv(max, newname, row.names = FALSE) #save as csv
}

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_max/", pattern="*.csv", full.names=TRUE)

final_max<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_max<-data.frame(park=park_codes, final_max)

#save
write.csv(final_max, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_max.csv", row.names = FALSE)

```