### 2020-12-06 Rcode/AQI_Alex.Rmd
AQI

---
title: "AQI"
author: "Alex Killion"
date: "12/4/2020"
output: html_document
---

```{r}
temp <- list.files("/Users/killion/Documents/GitHub/SmokeProject/Data/AQI", pattern="*.csv", full.names = TRUE) #county AQI by day 1980-2020
myfiles <- lapply(temp, read.csv)

all <- do.call(rbind,myfiles) #create on dataframe

library(dplyr)
library(lubridate)

all$Date <- ymd(all$Date) #reformat date

all$State.Code<-as.numeric(all$State.Code)
all$State.Code<-sprintf("%02d", all$State.Code)#convert state code to 2 digit to match .shp county file
all$County.Code<-sprintf("%03d", all$County.Code)#convert county code to 3 digit to match .shp county file
all$CODE<-paste0(all$State.Code, all$County.Code) #merge two codes to match to spatial data

median<-all %>% #get monthly means by county
  group_by(County=CODE, Year=year(Date), Month=month(Date)) %>%
  summarise(Monthly_Med=median(AQI))

library(sf)
library(sp)
library(rgeos)
#intersect parks with counties to get county ID for park
counties<-read_sf("/Users/killion/Documents/GitHub/SmokeProject/GIS/Counties/cb_2018_us_county_5m.shp")
parks<-read_sf("/Users/killion/Documents/GitHub/SmokeProject/GIS/NPS_poly/NPS_poly.shp")

parks_sp<-as_Spatial(parks) #convert filetype
parks2 <- gBuffer(parks_sp, byid=TRUE, width=0)#add buffer so polygons don't intersect self

parks3<-st_as_sf(parks2)#back convert file

counties<-st_transform(counties, crs=st_crs(parks)) #match projection

join<-st_join(parks3, counties, largest=TRUE) #add largest county id to each park

join_df<-as.data.frame(join)

#select only polygons designated as National Parks
join_df<-join_df[join_df$UNIT_TYPE=="National Park",]

#remove National Parks outside our Regions
join_df<-join_df[join_df$REGION=="PW"|join_df$REGION=="IM",]
join_df<-join_df[join_df$UNIT_CODE!="NPSA",] #Samoa
join_df<-join_df[join_df$STATE!="HI",]

join_df$County<-paste0(join_df$STATEFP, join_df$COUNTYFP) #create merged code to match AQI data

#keep only columns needed
parks_county<-join_df[,c(2,26)] #keep only park name and county/state ID

#join dataframes

final<-left_join(parks_county, median, by=c("County" = "County"))
write.csv(final, "/Users/killion/Documents/GitHub/SmokeProject/Data/AQI_Final_Data.csv")

```