### 2021-05-24 MarkdownDocuments/Manuscript_Real.Rmd
plots 'n writing

---
title: Breakpoint Models Show No Evidence of Recreational Adaptation to Increasing Wildfire Smoke in the American West
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken Hillis
- affiliation: Boise State University
  name: Laura? 
date: "`r format(Sys.time(), '%B %d, %Y')`"
keywords: Wildfire, Smoke, Recreation, Park Science
bibliography: [../SmokePaper.bib, ../OtherRefs.bib]
abstract: abstract text
fontsize: 12pt
link-citations: yes
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
header-includes:
- \usepackage{pdfpages}
- \usepackage[font={footnotesize}]{caption}
- \usepackage{bm}
- \usepackage{amsmath}

---
```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmarkdown)
library(bookdown)
library(formattable)
library(kableExtra)
library(dplyr)
library(magrittr)
library(prettydoc)
library(htmltools)
library(knitcitations)
library(devtools)
library(tidyverse)
library(scales)
library(extrafont)
```

```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
# Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)

## Caching code
knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")

## Plots
#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)

## Locate figures as close as possible to requested position (=code)
knitr::opts_chunk$set(fig.pos = 'H')

setwd("C:/Users/Matthewclark989/Documents/SmokeProject")
```

```{r plotzsrc, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
source("~/SmokeProject/Rcode/PaperPlotsFinal.R")
#source("~/SmokeProject/Rcode/PosteriorPlots.r")

```

# Introduction
Humanity must urgently adapt to our increasingly variable and hazardous climate [@IPCC]. Avoiding the worst-case projections of climate change will require adaptations at all levels of human society, from multinational governments to individuals [@IPCC]. Governments and individuals are aware of this, but still are failing to act [@steg_limiting_2018]. Knowledge of adaptive strategies and their benefits does not directly lead to adaptive behavior [@van_valkengoed_meta-analyses_2019]. While governments tend to implement climate adaptation policies in the wake of extreme events, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].

A particularly salient environmental effect of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016]. Because climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm. [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015]. Accompanying the increase in wildfires in the West, is a projected increase in wildfire smoke [@schroeder_validation_2008]. Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].

![image alt text. Righ-hand y axis shows the standardized values of smoke ](../Figures/smkplt.png)

Wildfire smoke is a complex mixture of gasses aerated by biomass combustion across the landscape [@naeher_woodsmoke_2007]. While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) is particularly harmful to human health [@deryugina_mortality_2019]. $\text{PM}_{2.5}$ Causes between 260,000 - 600,000 global deaths annually and significantly increasing community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].

Wildfires now account for 50% of all atmospheric $\text{PM}_{2.5}$ in some Western regions compared to less than 20% in 2010 [@burke_changing_2021]. As any level of ambient $\text{PM}_{2.5}$ increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].

Recent literature reviews on recreational behavior have found a dearth of research describing how individuals respond to low air quality, as well as research which gives predictions for how climate change will affect the recreation landscape [@otoole_climate_2019;@zajchowski_air_2019-1]. Additionally, a survey of federal land managers indicated that decision makers in areas such as national parks feel that they have a shortage of information describing how air quality affects the recreational behavior of their visitors [@zajchowski_air_2019].

It is of particular importance to understand how individuals are responding to wildfire smoke because it may provide a test case for how individuals will adapt to climate change generally. This article uses ambient smoke and visitation data collected from national parks in the Western and Intermountain United States from 1980-2019 to answer whether individuals are adapting their recreation behavior in response to increasing wildfire smoke. We hypothesize that the effect of ambient smoke on national park visitation may be nonlinear, only showing a measurable impact in the upper ranges of the observed smoke values. Threshold effects such as this are common in ecology and human behavioral sciences [@muradian2001ecological; @huggett2005concept; @macy1991chains]. Toms & Lesperance [-@toms2003piecewise] show that breakpoint models are an effective tool to account for these thresholds and gather inference about the effects of predictors before and after in natural systems. Here we develop a series of three hierarchical breakpoint models to determine if threshold effects are present in national park visitation response to ambient smoke and if so, what the effect of smoke is on visitation to national parks in the American West. The case study presented here aims to provide insights as to the overall trajectory of realized climate adaptation on the individual level.

# Methods
## Data Collection

### Visitation Data
Visitation data were obtained on the monthly scale through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/). These data are generally collected via car counters, permit information, and concessionaire reporting. Methods of collection vary from park to park. Unit specific information can be found on the National Park Service Visitor Use Statistics Portal. Although collection methods are inconsistent across the sample, all data was taken as reported. We retrieved monthly data for all national park units in the Intermountain and Pacific West regions for years 1980 through 2019. To reduce issues regarding park seasonality, we then subset our data to only the three month high season associated with each park. For example, only Summer (June, July, Aug) data was used for Yellowstone National Park.

### SmokeData
Alex to fill in this part

## Analyses

### Modeling
We first account for underlying change in national park visitation from 1980-2019 by developing a temporal autoregressive model which we validate and build upon to answer our main research question. This baseline autoregressive model uses the visitation data from the three previous years to predict the visitation to a given park in a given month. This formulation of an autoregressive model is commonly referred to as an AR(*k*) model, with this particular instance holding that *k* = 3 [@standev2021stancore]. We chose a *k* value of 3 as the added within-sample predictive value of increasing *k* diminished considerably for any greater value.

We formulated this baseline autoregressive model using a Bayesian hierarchical framework. This approach is appropriate here as it allows us to account between park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook]. For each park ($j$) we estimate a parameter value ($\bm{\nu1...k}$) representing the trend in visitation from one, two, and three years prior ($t-1,2,...k$), to the month we are estimating ($i$). Finally, we estimate both global ($\alpha0$) and park specific ($\bm{\beta0}_{j}$) intercepts to yield equation 1 below. We modeled these data using a negative binomial distribution to allow us to estimate an overdispersion parameter ($\phi$) rather than assuming the dispersion is equal to the mean.

$$ y_{ij} \sim NegBinom(\mu_{ij},\phi) $$

\begin{equation}
\text{log}(\mu_{ij}) = 
\alpha0+\bm{\beta0}_{j}+\bm{\nu1}_{j}*\text{vis}_{i,t-1}+ \\
\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3}  
\end{equation}

$$\bm{\beta0}_{j} \sim Norm(0,\sigma)$$

After validating that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate the if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold. To do this, we developed a series of three breakpoint models, allowing for three increasingly extreme breakpoints (i.e. critical thresholds). We standardized these smoke data, and thus the thresholds as well, by dividing by two standard deviations following the recommendation made by Gelman [-@gelman2008scaling]. A threshold of 1.0 can then be interpreted as two standard deviations above the mean value of smoke in the data. A visual comparison of the observed actual smoke values and standardized values can be seen in Fig. 1.

As shown conceptually in Fig. 2, we estimated the effect of smoke on visitation before ($\bm{\beta1}$) and after ($\bm{\beta2}$) each breakpoint (BP). Just as with the autoregressive terms, we estimated unique paramater values for each park ($j$) in a heierarchical framework. The complete model used for each of the three breakpoint values is then as seen in equation 2.

![image alt text](../Figures/bpplot2.png)

$$ y_{ij} \sim NegBinom(\mu_{ij},\phi) $$

\begin{equation}
\text{log}(\mu_{ij}) = 
\begin{cases} 
\alpha0+\bm{\beta0}_{j}+\bm{\beta1}_{j}*\text{smoke}_{i}+\bm{\nu1}_{j}*\text{vis}_{i,t-1}+ \\
\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3} & \text{if } \text{smoke}_{i} < \text{BP} \\ \\
\beta0+\bm{\beta0}_{j}+\bm{\beta1}_{j}*\text{BP}+(\text{smoke}_{i}-\text{BP})*\bm{\beta2}_{j}+  \\ \bm{\nu1}_{j}*\text{vis}_{i,t-1}+\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3}        & \text{if } \text{smoke}_{i} \geq \text{BP}
\end{cases}
\end{equation}

$$\bm{\beta0}_{j} \sim Norm(0,\sigma)$$

Following the recommendation given by Gelman et al. [-@gelman2008weakly] for producing stable, conservative estimates, we specified long-tailed regularizing priors for all autoregressive terms, the global intercept value, and the overdispersion parameter. We used normally distributed regularizing priors for estimating the variance parameter ($\sigma$) and the pre and post-breakpoint smoke effects to allow for more efficient sampling while generative within-sample predictions at high smoke values.

### Validation
Due to the considerable computational time (about 1 day) needed to run the program described above, we validated the results of both the autoregressive-only and breakpoint models using within-sample posterior predictive checks. All models were first inspected to ensure adequate mixing of Markov chains, $\hat{R}$ values below 1.05 for all parameter estimates, and the absence of divergent transitions after warm up.

We examined the within-sample predictive capacity of the autoregressive only model to be sure we adequately accounted for baseline changes in visitation before building on this model to test for the effects of smoke on visitation. This baseline model allowed us to account for approximately 40% of all variation in the data, with the bulk of the error occurring when actual visitation was very high or very low (Fig 3).

![image alt text](../Figures/ARVal.png)

# Results
While some between-park variation exists, our overall study shows no evidence for recreational visitor response to wildfire smoke in national parks in the American West. Figures 4 and 5 show the parameter estimates for the pre and post breakpoint smoke effects respectively. In all three breakpoint models, only two parks have 90% credible intervals that do not overlap zero for slope 1, both of which have opposite signs (Fig. 4). This trend is paralleled for slope 2 in the two lower breakpoint models and all 90% credibility intervals overlap 0 for slope 2 in the model with the most extreme breakpoint (Fig. 5).

Intuitively, the within-sample $R^2$ increases only slightly when accounting for smoke effects, from 0.40 to 0.41 compared to the autoregressive only model. This suggests that ambiant smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels. We then conclude that visitors are not altering recreation behavior in response to smoke and therefore no critical threshold for adaptation is detectable in these data.

![image alt text](../Figures/slpz1.png)

![image alt text](../Figures/slpz2.png)

# Discussion
We demonstrate that given the data available to us, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West. This result is troubling on both immediatly and in the long-term. We discuss earlier the public health dangers of wildfire smoke, and in particular $\text{PM}_{2.5}$.

We probably need national level policy to protect people [@rai_policy_2020]. There is huge variation in the way states react and plan for climate change [@hu_evaluating_2018].

because of the highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events [@dombeck_wildfire_2004; @collins_regional_2006].

\newpage

# References